### 2024-04-08 outputs/paper/paper.qmd
references updated, appendix formated

---
title: "TBD"
subtitle: "TBD"
author: 
  - Ziheng Zhong
thanks: "Code and data are available at: https://github.com/iJustinn/House_Price.git"
date: today
date-format: long
abstract: "TBD"
format: pdf
toc: true
toc-depth: 3
number-sections: true
bibliography: references.bib
---

```{r}
#| message: false
#| echo: false

# load packages
packages <- c("tidyverse", "boot", "broom.mixed", "collapse", "dataverse", "gutenbergr", "janitor", "knitr", "marginaleffects", "modelsummary", "rstanarm", "tidybayes", "loo", "kableExtra")

# Function to check and install missing packages
install_and_load <- function(packages) {
  for (package in packages) {
    if (!require(package, character.only = TRUE)) {
      install.packages(package)
      library(package, character.only = TRUE)
    }
  }
}

# Call the function with the list of packages
install_and_load(packages)
```

```{r}
#| message: false
#| echo: false

# read in data
house_prices <- read_csv("../../outputs/data/merged_house_price.csv")

state_house_price_by_year <- read_csv("../../outputs/data/state_house_price_by_year.csv")
state_1b_house_price_by_year <- read_csv("../../outputs/data/state_1b_house_price_by_year.csv")
state_2b_house_price_by_year <- read_csv("../../outputs/data/state_2b_house_price_by_year.csv")
state_3b_house_price_by_year <- read_csv("../../outputs/data/state_3b_house_price_by_year.csv")
state_4b_house_price_by_year <- read_csv("../../outputs/data/state_4b_house_price_by_year.csv")
state_5bplus_house_price_by_year <- read_csv("../../outputs/data/state_5bplus_house_price_by_year.csv")

state_house_price_by_month <- read_csv("../../outputs/data/state_house_price_by_month.csv")
state_1b_house_price_by_month <- read_csv("../../outputs/data/state_1b_house_price_by_month.csv")
state_2b_house_price_by_month <- read_csv("../../outputs/data/state_2b_house_price_by_month.csv")
state_3b_house_price_by_month <- read_csv("../../outputs/data/state_3b_house_price_by_month.csv")
state_4b_house_price_by_month <- read_csv("../../outputs/data/state_4b_house_price_by_month.csv")
state_5bplus_house_price_by_month <- read_csv("../../outputs/data/state_5bplus_house_price_by_month.csv")

```

# Introduction

# Data {#sec-data}

Data used in this paper was cleaned, processed and tested with the programming language R [@citeR]. Also with support of additional packages in R: `tidyverse` [@citeTidyverse], `ggplot2` [@citeGgplot], `janitor` [@citeJanitor], `readr` [@citeReadr], `knitr` [@citeKnitr], `modelsummary` [@citeModelsummary], `testthat` [@citeTestthat], `KableExtra` [@citeKableEx], `viridis` [@citeViridis], `lubridate` [@citeLubridate], `maps` [@citeMaps], `mgcv` [@citeMgcv].

## Source

## Method

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-sum
#| tbl-cap: Summary statistics of the California housing dataset
#| layout-ncol: 3
#| layout-nrow: 3

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-null-raw
#| tbl-cap: Count of missing values for each variable

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-null-clean
#| tbl-cap: Count of missing values for each variable after cleaning

```

# Results {#sec-res}

```{r}
#| include: false

```

## Data Trend

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-ocean_proximity_frequency
#| fig-cap: Frequency of Ocean Proximity

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-TBD_1
#| fig-cap: TBD

library(maps)
library(dplyr)
library(readr)
library(viridis)
library(ggplot2)

```

## Modeling

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-TBD_2
#| tbl-cap: TBD

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: fig-TBD_3
#| fig-cap: TBD

```

```{r}
#| message: false
#| echo: false
#| warning: false
#| fig-pos: 'h'
#| label: tbl-TBD_4
#| tbl-cap: TBD

```

# Discussion {#sec-dis}

## Demographic Shifts

## Health-related Behaviors

## Government Policies

## Environmental Changes

## Possible Improvements

# Conclusion {#sec-con}

\newpage

\appendix

# Appendix

## Datasheet

**Motivation**

**Composition**

**Collection process**

**Preprocessing/cleaning/labeling**

**Uses**

**Distribution**

**Maintenance**

\newpage

# References